\section{Reciprocally Inconsistent Distributed Edges}
\label{sec:db-corruption}

Suppose that the edge \texttt{(a)-[:\textcolor{green}{WROTE}]->(b)} is a distributed edge, with vertices \emph{a} and \emph{b} in servers $S_i$ and $S_j , j\neq i$,  respectively.
When a transaction writes this edge $ab$,
\begin{enumerate}
\item two writes are performed: reciprocal entries in the adjacency lists of both \emph{a} and \emph{b} are updated, and
\item write order is unconstrained: a transaction is equally likely to write \emph{a} then \emph{b} as it is to write \emph{b} then \emph{a}.
\end{enumerate}

Concurrent transactions $T_x$ and $T_y$ can interleave in the following three ways and each one is depicted in Fig \ref{fig:conf}:
\begin{enumerate}[(a)]
\item $T_x$ starts before $T_y$; it writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_y$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$ (see Fig \ref{fig:a}). The net effect is:  $T_x \rightarrow T_y$ at $S_i$ and at  $T_y \rightarrow T_x$ at $S_j$, where  $T \rightarrow T'$ at $S$ denotes that $T$ \emph{precedes} $T'$ at server $S$.
\item Same as the previous case, except that $T_y$ starts earlier than $T_x$ (see Fig \ref{fig:b}).
\item Same net effect as in previous two cases, except that both $T_x$ and $T_y$ start their first writes at $S_i$, $T_x \rightarrow T_y$, but $T_y$ overtakes $T_x$ in reaching $S_j$ where $T_y \rightarrow T_x$ (see Fig \ref{fig:c}).
\end{enumerate}

We could envisage three more corresponding cases (a') - (c') where the roles of $T_x$ and $T_y$ in cases (a) - (c) are simply interchanged; e.g., in case (a'), $T_y$ starts before $T_x$, writes $a$ at $S_i$ first and then proceeds to $S_j$ across the network; $T_x$ operates the other way round, beginning with $S_j$ and proceeding to $S_i$. Thus, there are only 6 ways concurrent $T_x$ and $T_y$ can interleave. The arguments we make based on cases (a) - (c) of Fig \ref{fig:conf} equally apply, by symmetry, to cases (a') - (c') and so, for brevity, we will not consider the latter.

At the end of each case in Fig \ref{fig:conf}, the last update on $a$ is by $T_y$ and that on $b$ is by $T_x$. Unless updates of $T_x$ and $T_y$ are commutative, $ab$ cannot be reciprocally consistent. As an example, suppose that $T_x$ deletes the \emph{wrote} edge while $T_y$ concurrently appends a property \emph{year}:
\input{cypher/delete-edge.tex}
\input{figures/hc}
The interleaving patterns depicted in Fig \ref{fig:conf} leave $ab$ reciprocally inconsistent, as shown in Fig \ref{hc}.

When $T_x$ and $T_y$ do not interleave,  either $T_y \rightarrow T_x$ or $T_x \rightarrow T_y$ holds at both servers $S_i$ and $S_j$; in that case, the last update on \emph{both} $a$ and $b$ would be by either $T_x$ or $T_y$, respectively. That is, when transactions update at both ends of $ab$  in \emph{some} arbitrarily chosen but identical order, $ab$ is left  reciprocally consistent after each transaction's update.

When interleaving updates leave $ab$ reciprocally inconsistent, $ab$ can be said to have become \emph{half-corrupted} because if $T_y \rightarrow T_x$ is the chosen order between $T_y$ and $T_x$, then the edge pointer in $b$ of $ab$ is in error; otherwise, the edge pointer in $a$ is erroneous. Thus, a reciprocally inconsistent edge $ab$ certainly has a corrupt half but the question of which half is corrupt has been left open.

Suppose that a future transaction $T_w$ \emph{first} reads the edge pointer of reciprocally in consistent $ab$, say, at vertex $a$. (Note that when $T_w$ reads any edge, it does not check for reciprocally consistency). At that moment, $T_w$ (implicitly) chooses the order $T_x \rightarrow T_y$ and thereby invalidates the other order $T_y \rightarrow T_x$ that prevails at vertex $b$. Thus, from that moment onward, the $b$ end of edge $ab$ becomes the corrupt end.

If no transaction ever reads the edge pointer at vertex $b$, then the order $T_x \rightarrow T_y$ effectively prevails and the half-corruption of $ab$ remains invisible to the rest of the database. However, if $T_z$ is to subsequently read the edge pointer at $b$ and write another edge based on what it read, i.e., The Hobbit has unknown author, then it is introducing updates not consistent with what $T_w$ read earlier; it thus introduces \emph{semantic corruption} into the database. Further writes based on reading semantically corrupt data also spread corruption.

A database is said to be \emph{operationally corrupt} when a significant proportion of its data records are in a semantically corrupt state. As stated earlier, we had shown that 10\%  of a distributed graph database can become semantically corrupt well within the system lifetime itself (\cite{Ezhilchelvan2018}, \cite{Webber2019}).

Two more relevant remarks on past works:
a half-corrupted edge is due to a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}) in the context of distributed graph databases. If the database provides the ANSI isolation level \emph{Read Uncommitted}, it will identically order the writes of concurrent transactions and this would prevent all interleaving patterns shown in Fig \ref{fig:conf} and thus avert half-corruption altogether.
